---
title: "Assessing the Impact of NPIs on Consumer Mobility Patterns and COVID-19 Transmission in the US"
author: "Joseph Zuccarelli, Laura Seaman, Kevin Rader"
date: "`r Sys.Date()`"
output:
  html_document: default

---

```{r setup, include=FALSE}
# Define necessary libraries
library(tidyr)
library(lubridate)
library(table1)
library(knitr)
library(dplyr)
library(lme4)
library(ggplot2)
library(cowplot)
library(scales)
library(reshape2)
library(broom.mixed)
library(zoo)
library(afex)
library(nlme)
```

In the following markdown document, we carry out all COVID-19 modeling using linear mixed effects models. 

## COVID-19 Modeling

First, we must read in our two datasets.
```{r}
# Read in data sets
MA_data <- read.csv("/Users/jqz0816/Documents/MDPI/Data/MA_data.csv")
MI_data <- read.csv("/Users/jqz0816/Documents/MDPI/Data/MI_data.csv")

# Format date/indicator features properly
MI_data$date <- as.Date(MI_data$date)
MA_data$date <- as.Date(MA_data$date)
MI_data$date_factor <- as.factor(MI_data$date)
MA_data$date_factor <- as.factor(MA_data$date)
MA_data$work_closing <- as.factor(MA_data$work_closing)
MA_data$gather_restrict <- as.factor(MA_data$gather_restrict)
MA_data$stay_home_reqs <- as.factor(MA_data$stay_home_reqs)
MI_data$work_closing <- as.factor(MI_data$work_closing)
MI_data$gather_restrict <- as.factor(MI_data$gather_restrict)
MI_data$stay_home_reqs <- as.factor(MI_data$stay_home_reqs)

# Group data by date
MA_data_sample <- MA_data %>%
  group_by(date) %>%
  summarise(num_transactions_tot = sum(num_transactions),
            num_businesses_tot = sum(num_unique_businesses),
            num_zips_tot = sum(num_unique_zips),
            num_morning_tot = sum(num_morning),
            num_afternoon_tot = sum(num_afternoon),
            num_evening_tot = sum(num_evening),
            num_night_tot = sum(num_night),
            num_grocery_tot = sum(num_grocery),
            num_convenience_tot = sum(num_convenience),
            num_health_tot = sum(num_health),
            num_hotel_tot = sum(num_hotels),
            num_restaurant_tot = sum(num_restaurants))
MI_data_sample <- MI_data %>%
  group_by(date) %>%
  summarise(num_transactions_tot = sum(num_transactions),
            num_businesses_tot = sum(num_unique_businesses),
            num_zips_tot = sum(num_unique_zips),
            num_morning_tot = sum(num_morning),
            num_afternoon_tot = sum(num_afternoon),
            num_evening_tot = sum(num_evening),
            num_night_tot = sum(num_night),
            num_grocery_tot = sum(num_grocery),
            num_convenience_tot = sum(num_convenience),
            num_health_tot = sum(num_health),
            num_hotel_tot = sum(num_hotels),
            num_restaurant_tot = sum(num_restaurants))

# Mutate additional sample-level features
MA_data_idv <- MA_data %>% 
  filter(uuid=="0007561724B170454D13B2F6CE227C3D77F116951F5A75BC258349DD18AD0733") %>%
  select(day,weekend,emergency,work_closing,gather_restrict,stay_home_reqs,
         case_count,case_rate,test_count,test_rate,date_factor) 
MI_data_idv <- MI_data %>% 
  filter(uuid=="000919F1AAA2A62098038034B268628BCAE1157EBC627042B45BFD7AF38A22C6") %>%
  select(day,weekend,emergency,work_closing,gather_restrict,stay_home_reqs,
         case_count,case_rate,test_count,test_rate,date_factor) 
MA_data_sample <- cbind(MA_data_sample,MA_data_idv)
MI_data_sample <- cbind(MI_data_sample,MI_data_idv)
MA_data_sample$state <- as.factor('Boston, MA')
MI_data_sample$state <- as.factor('Ann Arbor, MI')
```

Now, let's mutate the primary two features required for modeling: COVID-19 case growth rates and lagged delta transactions.
```{r}
# Create exp growth rate features
MA_data_sample <- MA_data_sample %>%
  mutate(case_growth_rate = 100*(log(case_count+1) - log(lag(case_count+1))))
MI_data_sample <- MI_data_sample %>%
  mutate(case_growth_rate = 100*(log(case_count+1) - log(lag(case_count+1))))

# Calculate pre-pandemic baseline transaction counts
MA_baseline <- MA_data_sample %>% 
  filter(date < "2020-03-01") %>%
  group_by(day) %>%
  summarise(baseline_transactions = median(num_transactions_tot))
MI_baseline <- MI_data_sample %>% 
  filter(date < "2020-03-01") %>%
  group_by(day) %>%
  summarise(baseline_transactions = median(num_transactions_tot))

# Merge baseline transaction feature
MA_data_sample <- merge(MA_data_sample,MA_baseline,by='day') %>%
  arrange(date) %>%
  mutate(delta_transactions = 100*((num_transactions_tot-baseline_transactions)/baseline_transactions))
MI_data_sample <- merge(MI_data_sample,MI_baseline,by='day') %>%
  arrange(date) %>%
  mutate(delta_transactions = 100*((num_transactions_tot-baseline_transactions)/baseline_transactions))
data_sample_plot <- rbind(MI_data_sample,MA_data_sample)
data_sample_plot$state <- factor(data_sample_plot$state, levels = c('Boston, MA',
                                                          'Ann Arbor, MI'))

# Mutate lagged transaction features
MA_data_emergency <- MA_data_sample %>%
  mutate(delta_trans_lag18 = lag(delta_transactions, n = 18),
         delta_trans_lag16 = lag(delta_transactions, n = 16),
         delta_trans_lag14 = lag(delta_transactions, n = 14),
         delta_trans_lag12 = lag(delta_transactions, n = 12),
         delta_trans_lag10 = lag(delta_transactions, n = 10),
         delta_trans_lag8 = lag(delta_transactions, n = 8),
         delta_trans_lag6 = lag(delta_transactions, n = 6)) %>%
  filter(emergency==1)
MI_data_emergency <- MI_data_sample %>%
  mutate(delta_trans_lag18 = lag(delta_transactions, n = 18),
         delta_trans_lag16 = lag(delta_transactions, n = 16),
         delta_trans_lag14 = lag(delta_transactions, n = 14),
         delta_trans_lag12 = lag(delta_transactions, n = 12),
         delta_trans_lag10 = lag(delta_transactions, n = 10),
         delta_trans_lag8 = lag(delta_transactions, n = 8),
         delta_trans_lag6 = lag(delta_transactions, n = 6)) %>%
  filter(emergency==1)

# Combine state DFs
data_emergency <- rbind(MA_data_emergency,MI_data_emergency)
```

Prior to performing any modeling, it is necessary to visualize the distributions of our response variables.
```{r fig.align="center",fig.width = 10,warning=FALSE}
# Histogram: Distribution of Case Growth Rate (MA)
MA_plot <- MA_data_emergency %>%
  ggplot(aes(x=case_growth_rate)) +
  geom_histogram()+
  labs(x="Case Growth Rate",y="Count",title="Distribution of Case Growth Rate (MA)") 

# Histogram: Distribution of Case Growth Rate (MA)
MI_plot <- MI_data_emergency %>%
  ggplot(aes(x=case_growth_rate)) +
  geom_histogram()+
  labs(x="Case Growth Rate",y="Count",title="Distribution of Case Growth Rate (MI)") 

# Subplots
plot_grid(MA_plot,MI_plot)
```
Note that for both samples the response variable appears to be normally distributed, suggesting that linear models are appropriate.

Let's also visualize our primary explanatory variable, $\Delta$ transactions.
```{r fig.align="center",warning=FALSE,message=FALSE,cache=TRUE}
# Plot: Sample Average Daily Businesses by Date
emergency <- as.Date('2020-03-13')
data_sample_plot %>%
  mutate(avg_delta_trans = rollmean(delta_transactions, k=7, fill=NA, align='right')) %>%
  filter(date >= '2020-03-01') %>%
  ggplot(aes(x=date,y=avg_delta_trans,color=state)) +
  geom_line() +
  geom_text(aes(x=ymd("2020-03-13"),y=(-14),label='NATIONAL'),color='red')+
  geom_text(aes(x=ymd("2020-03-13"),y=(-18),label='EMERGENCY'),color='red')+
  annotate(geom = "vline",
           x = c(emergency),
           xintercept = c(emergency),
           linetype = c("dashed"),
           color = c("red")) +
  geom_hline(yintercept = 0, linetype="dashed") +
  labs(x='Date',y='\u0394 Transactions (% Chg vs. Pre-Pandemic Baseline)', color="Sample",
       title='\u0394 Transactions by Date')
```
Finally, let's begin modeling. Formally, our COVID-19 models are defined as follows:

\begin{equation}
   case \: growth \: rate_{ij} = \beta_0 + \beta_{1} \Delta transactions_{ijk} + \beta_{2} test \: rate_{ij} + \beta_{3} weekend_{i} + \gamma_{0j} .    
\end{equation}

The response variable, $case \: growth \: rate_{ij}$, denotes the confirmed COVID-19 case growth rate observed on date $i$ for county $j$. The explanatory variable of interest, $\Delta transactions_{ijk}$, represents the $k$-day lagged percentage of change in each sample's daily total POS transaction count compared to baseline. We controlled for the state-level COVID-19 testing rate (per 100,000 population) and seasonal patterns induced by the day of the week (see variables $test \: rate_{ij}$ and $weekend_i$). Similarly, we adjusted for county-level effects with the term $\gamma_{0j}$, which represents the random intercept for county $j$. We fit each model using a first-order autoregressive covariance structure with homogeneous variances to account for the time series nature of our data.

**Lag: 08 days**
```{r,results='hide',warning=FALSE}
# Model: Case Growth Rate ~ Delta Transaction Rate (Lag 08 days)
model_8days <- lme(case_growth_rate ~ delta_trans_lag8 + test_rate + weekend,
               random = ~1 |state,
             correlation = corAR1(form = ~date),
             data = data_emergency)
```

```{r,results='hide',warning=FALSE}
# Model summary
summary(model_8days)
```

```{r}
# Model output
tidy(model_8days,effects = 'fixed',conf.int = TRUE)
```

```{r}
# Random Coefficients
coef(model_8days)
```

**Lag: 10 days**
```{r,results='hide',warning=FALSE}
# Model: Case Growth Rate ~ Delta Transaction Rate (Lag 10 days)
model_10days <- lme(case_growth_rate ~ delta_trans_lag10 + test_rate + weekend,
               random = ~1 |state,
             correlation = corAR1(form = ~date),
             data = data_emergency)
```

```{r,results='hide',warning=FALSE}
# Model summary
summary(model_10days)
```

```{r}
# Model output
tidy(model_10days,effects = 'fixed',conf.int = TRUE)
```

```{r}
# Random Coefficients
coef(model_10days)
```

**Lag: 12 days**
```{r,results='hide',warning=FALSE}
# Model: Case Growth Rate ~ Delta Transaction Rate (Lag 12 days)
model_12days <- lme(case_growth_rate ~ delta_trans_lag12 + test_rate + weekend,
               random = ~1 |state,
             correlation = corAR1(form = ~date),
             data = data_emergency)
```

```{r,results='hide',warning=FALSE}
# Model summary
summary(model_12days)
```

```{r}
# Model output
tidy(model_12days,effects = 'fixed',conf.int = TRUE)
```

```{r}
# Random Coefficients
coef(model_12days)
```

**Lag: 14 days**
```{r,results='hide',warning=FALSE}
# Model: Case Growth Rate ~ Delta Transaction Rate (Lag 14 days)
model_14days <- lme(case_growth_rate ~ delta_trans_lag8 + test_rate + weekend,
               random = ~1 |state,
             correlation = corAR1(form = ~date),
             data = data_emergency)
```

```{r,results='hide',warning=FALSE}
# Model summary
summary(model_14days)
```

```{r}
# Model output
tidy(model_14days,effects = 'fixed',conf.int = TRUE)
```

```{r}
# Random Coefficients
coef(model_14days)
```
