---
title: "Assessing the Impact of NPIs on Consumer Mobility Patterns and COVID-19 Transmission in the US"
author: "Joseph Zuccarelli, Laura Seaman, Kevin Rader"
date: "`r Sys.Date()`"
output:
  html_document: default

---

```{r setup, include=FALSE}
# Define necessary libraries
library(tidyr)
library(lubridate)
library(table1)
library(knitr)
library(dplyr)
library(lme4)
library(ggplot2)
library(cowplot)
library(scales)
library(reshape2)
library(broom.mixed)
library(zoo)
library(afex)
library(iccCounts)
library(glmmTMB)
library(performance)
```

In the following markdown document, we carry out all mobility modeling using generalized linear mixed effects models. 

## Mobility Modeling

First, we must read in our two datasets.
```{r}
# Read in data sets
MA_data <- read.csv("/Users/jqz0816/Documents/MDPI/Data/MA_data.csv")
MI_data <- read.csv("/Users/jqz0816/Documents/MDPI/Data/MI_data.csv")

# Format date/indicator features properly
MI_data$date <- as.Date(MI_data$date)
MA_data$date <- as.Date(MA_data$date)
MI_data$date_factor <- as.factor(MI_data$date)
MA_data$date_factor <- as.factor(MA_data$date)
MA_data$work_closing <- as.factor(MA_data$work_closing)
MA_data$gather_restrict <- as.factor(MA_data$gather_restrict)
MA_data$stay_home_reqs <- as.factor(MA_data$stay_home_reqs)
MI_data$work_closing <- as.factor(MI_data$work_closing)
MI_data$gather_restrict <- as.factor(MI_data$gather_restrict)
MI_data$stay_home_reqs <- as.factor(MI_data$stay_home_reqs)
```

Prior to performing any modeling, it is necessary to visualize the distributions of our variables.
```{r fig.align="center",warning=FALSE}
# Histogram: Daily Transactions Count (MA Jan - May 2020)
MA_plot <- MA_data %>%
  ggplot(aes(x=num_transactions)) +
  geom_histogram()+
  labs(x="Number of Daily Transactions",y="Count",title="Daily Transactions Count (MA)") 

# Histogram: Daily Transactions Count (MI Jan - May 2020)
MI_plot <- MA_data %>%
  ggplot(aes(x=num_transactions)) +
  geom_histogram()+
  labs(x="Number of Daily Transactions",y="Count",title="Daily Transactions Count (MI)") 

# Subplots
plot_grid(MA_plot,MI_plot)
```
The plots above indicate that our data is heavily right-skewed; therefore, Poisson models are most appropriate. We will also check for over-dispersion, in which case negative binomial models would be necessary. Formally, our consumer mobility models are defined as follows:

\begin{equation} 
  \log(transactions_{ij})  = \beta_0 + \beta NPI_{i} + \beta weekend_{i} + \gamma_{0j}. 
\end{equation}

The response variable, $transactions_{ij}$, denotes the POS transaction count observed on date $i$ for individual $j$. The indicator variable, $NPI_{i}$, represents one of the four interventions of interest in our study. Note that we treated each NPI as a categorical variable with multiple levels; therefore, fitting each separate model with four different NPI variables resulted in multiple $\beta$ coefficients due to one hot encoding. We controlled for seasonal patterns induced by the day of the week with the variable $weekend_{i}$. Similarly, we adjusted for individual-level effects with the term $\gamma_{0j}$, which represents the random intercept for individual $j$. We fit each model using a first-order autoregressive covariance structure with homogeneous variances to account for the time series nature of our data. 

**Massachusetts**

**NPI: National Emergency**
```{r,results='hide',warning=FALSE}
# Model: Daily Number of Transactions ~ National Emergency (MA)
modelMA <- glmmTMB(num_transactions ~ emergency + weekend +  (1|uuid) +
                  + ar1(date_factor + 0 |uuid),
                   family = poisson(),
                   data=MA_data)
```

```{r}
# Overdispersion test
check_overdispersion(modelMA)
```

```{r}
# Model summary
summary(modelMA)
```

```{r}
# Fixed Effects
tidy(modelMA,exponentiate = TRUE,effects = 'fixed',conf.int = TRUE)
```

```{r}
# ICC
rand_model <- icc_counts(MA_data,y='num_transactions',id='uuid',fam="poisson")
rand_model$ICC
```

**NPI: Workplace Closing**
```{r,results='hide',warning=FALSE}
# Model: Daily Number of Transactions ~ Workplace Closing (MA)
modelMA <- glmmTMB(num_transactions ~ work_closing + weekend + (1 | uuid) +
                   ar1(date_factor + 0 |uuid),
                   family = poisson(),
                   data=MA_data)
```

```{r}
# Overdispersion test
check_overdispersion(modelMA)
```

```{r}
# Model summary
summary(modelMA)
```

```{r}
# Model output
tidy(modelMA,exponentiate = TRUE,effects = 'fixed',conf.int = TRUE)
```

```{r}
# ICC
rand_model <- icc_counts(MA_data,y='num_transactions',id='uuid',fam="poisson")
rand_model$ICC
```

**NPI: Stay-at-Home Requirements**
```{r,results='hide',warning=FALSE}
# Model: Daily Number of Transactions ~ Stay-at-Home Requirements (MA)
modelMA <- glmmTMB(num_transactions ~ stay_home_reqs + weekend + (1 |uuid) +
                   ar1(date_factor + 0 |uuid),
                   family = poisson(),
                   data=MA_data)
```

```{r}
# Overdispersion test
check_overdispersion(modelMA)
```

```{r}
# Model summary
summary(modelMA)
```

```{r}
# Model output
tidy(modelMA,exponentiate = TRUE,effects = 'fixed',conf.int = TRUE)
```

```{r}
# ICC
rand_model <- icc_counts(MA_data,y='num_transactions',id='uuid',fam="poisson")
rand_model$ICC
```

**NPI: Gathering Restrictions**
```{r,results='hide',warning=FALSE}
# Model: Daily Number of Transactions ~ Gathering Restrictions (MA)
modelMA <- glmmTMB(num_transactions ~ gather_restrict + weekend + (1|uuid) +
                   ar1(date_factor + 0 |uuid),
                   family = poisson(),
                   data=MA_data)
```

```{r}
# Overdispersion test
check_overdispersion(modelMA)
```

```{r}
# Model summary
summary(modelMA)
```

```{r}
# Model output
tidy(modelMA,exponentiate = TRUE,effects = 'fixed',conf.int = TRUE)
```

```{r}
# ICC
rand_model <- icc_counts(MA_data,y='num_transactions',id='uuid',fam="poisson")
rand_model$ICC
```

**Michigan**

**NPI: National Emergency**
```{r,results='hide',warning=FALSE}
# Model: Daily Number of Transactions ~ National Emergency (MI)
modelMI <- glmmTMB(num_transactions ~ emergency + weekend + (1|uuid) +
                  + ar1(date_factor + 0 |uuid),
                   family = poisson(),
                   data=MI_data)
```

```{r}
# Overdispersion test
check_overdispersion(modelMI)
```

```{r}
# Model summary
summary(modelMI)
```

```{r}
# Model output
tidy(modelMI,exponentiate = TRUE,effects = 'fixed',conf.int = TRUE)
```

```{r}
# ICC
rand_model <- icc_counts(MI_data,y='num_transactions',id='uuid',fam="poisson")
rand_model$ICC
```

**NPI: Workplace Closing**
```{r,results='hide',warning=FALSE}
# Model: Daily Number of Transactions ~ Workplace Closing (MI)
modelMI <- glmmTMB(num_transactions ~ work_closing + weekend + (1 |uuid) +
                   ar1(date_factor + 0 |uuid),
                   family = poisson(),
                   data=MI_data)
```

```{r}
# Overdispersion test
check_overdispersion(modelMI)
```

```{r}
# Model summary
summary(modelMI)
```

```{r}
# Model output
tidy(modelMI,exponentiate = TRUE,effects = 'fixed',conf.int = TRUE)
```

```{r}
# ICC
rand_model <- icc_counts(MI_data,y='num_transactions',id='uuid',fam="poisson")
rand_model$ICC
```

**NPI: Stay-at-Home Requirements**
```{r,results='hide',warning=FALSE}
# Model: Daily Number of Transactions ~ Stay-at-Home Requirements (MI)
modelMI <- glmmTMB(num_transactions ~ stay_home_reqs + weekend + (1 |uuid) +
                   ar1(date_factor + 0 |uuid),
                   family = poisson(),
                   data=MI_data)
```

```{r}
# Overdispersion test
check_overdispersion(modelMI)
```

```{r}
# Model summary
summary(modelMI)
```

```{r}
# Model output
tidy(modelMI,exponentiate = TRUE,effects = 'fixed',conf.int = TRUE)
```

```{r}
# ICC
rand_model <- icc_counts(MI_data,y='num_transactions',id='uuid',fam="poisson")
rand_model$ICC
```

**NPI: Gathering Restrictions**
```{r,results='hide',warning=FALSE}
# Model: Daily Number of Transactions ~ Gathering Restrictions (MI)
modelMI <- glmmTMB(num_transactions ~ gather_restrict + weekend + (1|uuid) +
                   ar1(date_factor + 0 |uuid),
                   family = poisson(),
                   data=MI_data)
```

```{r}
# Overdispersion test
check_overdispersion(modelMI)
```

```{r}
# Model summary
summary(modelMI)
```

```{r}
# Model output
tidy(modelMI,exponentiate = TRUE,effects = 'fixed',conf.int = TRUE)
```

```{r}
# ICC
rand_model <- icc_counts(MI_data,y='num_transactions',id='uuid',fam="poisson")
rand_model$ICC
```











